---
title: "COMPSCIX 415.2 Homework 1"
author: "Chunduri SriKeerthi Chandra"
date: "Jun 12th 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/skcchunduri/compscix-415-2-assignments.git](your URL)


```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
data("WorldCities")
```

#### QUESTION 1: How many observations and variables are there in this data set? What are
#### some of the variable names? Type up your answer in your document using complete sentences.

WorldCities dataset captures **`r nrow(WorldCities)`** observations made for **`r ncol(WorldCities)`** variables ( `r names(WorldCities)` ). These variables capture different aspects of each city that help locate the city and know its population.

```{r}
WorldCities <- head(WorldCities, 200)
country_col <- WorldCities$country
```

Countries from sample dataset [ `r unique(country_col)` ]. Different regions that these countries belong to are :

#### QUESTION 2: There is a column called region in the data set. Can you extract this and show
#### only the unique values?

```{r}
WorldCities %>% select(region) %>% unique()
```

#### QUESTION 3: Can you extract and show the unique entries from the country column in
#### WorldCities using one line of code and two %>% operators?

A different version to select countries is:
```{r}
WorldCities %>% select(country) %>% unique()
```

### Histogtam
```{r}
WorldCities %>% ggplot(aes(x = population)) +
geom_histogram() +
xlab('City population')
```

#### QUESTION 4: Make one more histogram of the population and add a new x-label, a new ylabel,
#### a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().

My Histogram using "classic dark-on-light ggplot2 theme"
```{r}
WorldCities %>% ggplot(aes(x = population)) +geom_histogram() +
  theme_bw() + xlab('population') + ylab('frequency') + ggtitle('classic dark-on-light ggplot2 theme')
```
