---
title: "“COMPSCIX 415.2 Homework 4"
author: "Chunduri SriKeerthi Chandra"
date: "Jul 3rd 2018"
output: 
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initilize, include=FALSE}
library(tidyverse)
library(nycflights13)
```

##**Data munging and data wrangling**

###5.6.7 Exercises

***Q:** Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).*

```{r}
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

Total number of flights not canceled for a destination can also be computed as
```{r}
not_cancelled %>% group_by(dest) %>% summarise(n())
```

Totla distance each flight flew can be computed as:

```{r}
not_cancelled %>% group_by(tailnum) %>% summarise(n = sum(distance))
```

***Q:** Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?*


```{r}

mutated_flights <- flights %>% group_by(year, month, day)  %>% mutate(canceled = (is.na(dep_delay) & is.na(arr_delay)))

```

Total number of canceled flights per day are:
```{r}
mutated_flights %>% summarise(canceled = sum(canceled))
```

Comparing average departure delay Vs average flight cancelation:
```{r}
(data <- mutated_flights %>% summarise( count = n(), canceled = mean(canceled, na.rm = TRUE), delay = mean(dep_delay, na.rm = TRUE)) %>% filter(count > 50))

data %>% ggplot() + geom_point(aes(x = delay, y = canceled, color = count), position = "jitter")
```


From the plot above, It looks like more number of flights are canceled within the first 20 min of delay. Flight delays and cancelations are not related. 

***Q:** What does the sort argument to count() do. When might you use it?*

When *sort* is set to *TRUE* it will arrange the rows in the desending order of n, similar to arrange(desc(n)) verb.

###10.5 Exercises

***Q:** 1 How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).*
We can test to see if the object is a tibble like:
```{r}
is.tibble(flights)
```

Or we can display the object to tell if its a tibble. Tibbles have a refined print method that shows only the first 10 rows:
```{r}
flights
```

Or we can check the class of the object
```{r}
class(flights)
```


***Q:**2 Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?*
*df <- data.frame(abc = 1, xyz = "a")*
*df$x*
*df[, "xyz"]*
*df[, c("abc", "xyz")]*

Tibbles are strict when accessing columns using *$* or *[[]]*. Whereas *data.frames* match the nearest column name, as in this case *data.frame* doesn't have column **x**, but still displays *xyz* value.
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

Which is not the case with tibbel:
```{r}
df <- tibble(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

And the tibble displays the column type which is more informative compared to nothing on *data.frames*.

***Q:**3 If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?*

We can use **[[]]** or **$** to read a variable contents from tibble.

***Q:**6 What option controls how many additional column names are printed at the footer of a tibble?*

*print()* function can be used to control the number of rows and number of columns to be displayed.
*print()* takes variables *n* to control number of rown and *width* to control number of columns.

###12.3.3 Exercises

***Q:**2 Why does this code fail?*
*table4a %>%*
  *gather(1999, 2000, key = "year", value = "cases")*

Column names 1999 and 2000 doesn't start with a letter, since they are non-syntactic names we have to encode them in backticks otherwise R takes them as range.

correct way to transfer:
```{r}
table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
```

***Q:**3 Why does spreading this tibble fail? How could you add a new column to fix the problem?*
```{r error=TRUE}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

people %>% spread(key = key, value = value)
```

observations on row 1 and row 3 are duplicate for age, hence spread fails with *Error: Duplicate identifiers for rows (1, 3)* message.

**Note**: Not quite sure how to get around this issue. Not answered.

***Q:**4 Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?*
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

We have to gather the male and female in to a colum. 

```{r}
preg %>% gather(male, female, key=gender, value=count)
```

###12.4.3 Exercises

***Q:**1 What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.*

*extra* argument can be used to control what happens when a split generates more than what is required number of values. And *fill* argument can be used to control how *NA* placed when a split generates less than the required number values.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  #Drops the extra value silently
  separate(x, c("one", "two", "three"), extra = "drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  #Fills NA on the left instead of on the right(default)
  separate(x, c("one", "two", "three"), fill = "left")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  #Only splits at most lenght(into), that means "d,e,f,g" is split in to "d" and "e" and "f,g"
  separate(x, c("one", "two", "three"), extra = "merge")

```

***Q:**2 Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?*

*remove* argument can be used to retain the original column in the output data frame. The default value is *TRUE* which removes the original column, but when it is set to *FALSE* output data frame retains the original column.

Example:

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  #Drops the extra value silently
  separate(x, c("one", "two", "three"), extra = "drop", remove = FALSE)
```

### Data Import Exercises

Follow these steps:
- Download the baby_names.txt file from Canvas which is in the Homework 4 assignment section.
- Load this file into R correctly and take a glimpse of the output.
- Export this file as a csv file and call it ‘baby_names.csv’.
- Reload the baby_names.csv file and take another glimpse.
- Show all of your code and the output.

```{r}
#Loading baby_names.txt
filePath <- "/Users/srchundu/UCBExtension/DataScienceIntro/compscix-415-2-assignments/Lab4/baby_names.txt"

#File is a "|" seperated
fileData <- read_delim(file = filePath, delim = "|")

#Glimpse at file content
fileData

# Write file contents back to CSV file
csvFilePath <- "/Users/srchundu/UCBExtension/DataScienceIntro/compscix-415-2-assignments/Lab4/baby_names.csv"
fileData %>% write_csv(path = csvFilePath)

# Read back baby names from CSV file
fileData2 <- read_csv(file = csvFilePath)

#Glimpse at file content
fileData2
```

### Challenge Question 5.6.7 Exercise

***Q:**5 Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))*

I couldn't derive the relation between bad airports and bad carriers

Carriers with worst delays:
```{r}
delayed_flights <- flights %>% filter((!is.na(dep_delay) & dep_delay > 0) | (!is.na(arr_delay) & arr_delay > 0));
delayed_flights %>% group_by(carrier) %>% summarise(total = n()) %>% arrange(desc(total))
```

Airports with worst delays:
```{r}
delayed_flights <- flights %>% filter((!is.na(dep_delay) & dep_delay > 0) | (!is.na(arr_delay) & arr_delay > 0))
delayed_flights %>% group_by(dest) %>% summarise(total = n()) %>% arrange(desc(total))
```

