---
title: "“COMPSCIX 415.2 Homework 3"
author: "Chunduri SriKeerthi Chandra"
date: "Jun 25th 2018"
output: 
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initilize, include=FALSE}
library(tidyverse)
library(nycflights13)
data(mpg)
```

##**Data Visualisation/Transformation**

###3.8.1 Exercises

***Q:** What is the problem with this plot? How could you improve it?*

The overlapping points are rounted and so it is hard to find how the data is distributed. We can fix it by adding jitter.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

***Q:** What parameters to geom_jitter() control the amount of jittering?*

*width* and *Height* parameters controll the amount of jittering in *geom_jitter()*.

***Q:** Compare and contrast geom_jitter() with geom_count()*

**NOT ANSWERED**

***Q:** What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.*

Default position object is dodge. 
```{r}
ggplot(data = mpg, aes(x = drv, y = hwy, colour = class)) +
  geom_boxplot()
```


###3.9.1 Exercises

***Q:** What does labs() do? Read the documentation.*

**labs()** help provide seld explanatory labels on x/y axis and for the title of the graph.

***Q:** What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?*

highway mileage and city mileage are proportional. As the city milage increases highway mileage increases. *coord_fixed* will enforce the specified ratio on the physical representation of data units on the axes. The default ratio = 1. *geom_abline* adds a reference lines to the plot.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

###4.4 Exercises

***Q:** Why does this code not work?*

The initialized variable is *my_variable*. R doesn't know *my_var1able*. **1** should be replaced by **i**.

***Q:** Tweak each of the following R commands so that they run correctly*

Correct commands : 
```{r}
library(tidyverse)
data("diamonds")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```


###5.2.4 Exercises

***Q:** Find all flights that *
  *1.Had an arrival delay of two or more hours*
```{r}
filter(flights, arr_delay >= 2*60)
```
  
  *2. Flew to Houston (IAH or HOU)*
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```
  *3. Were operated by United, American, or Delta*
```{r}
filter(flights, carrier %in% c("AA", "DL", "UA"))
```
  *4. Departed in summer (July, August, and September)*
```{r}
filter(flights, month %in% c(6, 7, 8))
```

  *5. Arrived more than two hours late, but didn’t leave late*
```{r}
filter(flights, dep_delay == 0, arr_delay > 2*60)
```
  
  *6. Were delayed by at least an hour, but made up over 30 minutes in flight*
```{r}
filter(flights, dep_delay >= 60, arr_delay == (dep_delay - 30))
```
  
  *7. Departed between midnight and 6am (inclusive)*
```{r}
filter(flights, dep_time <= 600)
```
  
  ***Q:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?*
```{r}
depMissing <- filter(flights, is.na(dep_time))
depMissing
```

flight data set has `r NROW(depMissing)` missing dep_time observations. It also has *'dep_delay'*, *'arr_time'*, *'arr_delay'* and *'air_time'*. These observations/rows represents scheduled flights that where cancelled.

***Q:** Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA \* 0 is a tricky counterexample!)*

**NOT ANSWERED**

###  5.4.1 Exercises

***Q:** Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights. *
  1
```{r}
select(flights, dep_time, dep_delay, arr_time)
```
  2
```{r}
select(flights, dep_time, dep_delay:arr_time)
```
  3
```{r}
select(flights, -(year:day),-sched_dep_time,-(sched_arr_time:time_hour))
```

***Q:** What does the one_of() function do? Why might it be helpful in conjunction with this vector?*

*One_of* function helps select the available columns from the table. Using vector to store all the possible column names to select will help reuseing.
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```


